---
title: "R Notebook"
output: html_notebook
---

CCA con VEGAN

```{r}
setwd("~/Dropbox/maestría")
practicas1<-read.table("CCAManejo.csv", sep=",", header=TRUE) 
familias1<-read.table("CCAfamilias.csv", sep=",", header=TRUE)
practicas <- practicas1[, 3:24]
familias <- familias1[, 2:26]
```
```{r}
#fijándome sólo en las más abundantes de los guilds herbivoros, depredadres, polifagos y detritivoros(sapro), o sea, quito todo menos Chrysomelidae, Curculionidae, Coccinellidae, Cleridae, Carabidae, Anthicidae, Phalacridae, Nitidulidae, Aderidae
abundance.matrix=subset(familias, select=c(Chrysomellidae, Curculionidae, Coccinellidae, Cleridae, Carabidae, Nitidullidae, Anobiidae, Cantharidae, Phalacridae))
abundance.matrix
abundance.matrix<-data.matrix(abundance.matrix) #lo hago matriz

```
```{r}
#implement a square root transformation to dampen the influence of large counts on the results
abundance.matrix2 <- sqrt(abundance.matrix)
abundance.matrix2
```
First, we will apply correspondence analysis (CA) to the beetle abundance matrix and visualize the major axes of variation in beetle community composition. Next, we will perform a constrained CA, also known as canonical CA (CCA), bringing the management data into play as predictors of beetle community composition.


```{r}
my.ca <- cca(abundance.matrix2)
barplot(my.ca$CA$eig/my.ca$tot.chi, names.arg = 1:my.ca$CA$rank, cex.names = 0.5, ylab="Proportion of variance explained", xlab="CA axis")
# CA proportions of variance explained
head(my.ca$CA$eig/my.ca$CA$tot.chi)
```
The fact that the first 2 axes explain a combined ~50% of variation in the data indicates that beetle community composition is not easily reduced to just one or two axes of variation. Let’s look at a scatterplot of our data with respect to the first two CAs:


```{r}
plot(my.ca, scaling="symmetric", main="scaling = 3, both")
#text(my.ca,  labels= practicas1$manejo)
```
An alternative way to make more space on our plot is to simply display the sites and species on separate plots. Let’s also label sites to represent our typology, and increase the size of the labels.
```{r}
# 2 panels
layout(matrix(1:2,1,2))

# plot sites
plot(my.ca, display="sites", type="n", main="Sites", scaling="sites")
text(my.ca, display="sites", labels= practicas1$manejo,scaling="sites")

# plot species
plot(my.ca, display="species", type="n", ylab="", main="Families", scaling="species")
text(my.ca, display="species", col="black", scaling="species")
```
That’s better. The labels help reveal that the second CA axis reflects a management gradient, suggesting that management is an  important factor structuring beetle community composition (though not the most important one, as it is the second and not the first axis that reflects this). 

With the sites and species separated out, it is easier to see the differentiation among species as well. Given that CA2 appears to be an industrial-traditional gradient, we can see that some beetle species tend to be found in the traditional plots (e.g. Curculionidae, Nitidulidae, Cleridae), while others tend to be found in the industrialized ones (e.g. Carabidae). Species near the center of the graph do not show much differentiation among either regions or habitats.



Now we will bring in the management variables. Let’s make a subset of the data for the managemebt variables by substracting those who are too correlated >70% (number of varieties and monoculture, leving number of crops; fallow period leaving irrigation) and also those that contributed very scarcely to both PCA components 1 and 2 (stubble, comercial.agriculture, manure, subsistence.agriculture, organic.fertilizer, native.seeds, green.manure) - these will be our predictor variables:

```{r}
# predictor variables
fixed=subset(practicas, select=-c(number.varieties, monoculture, fallow.period, stubble, comercial.agr, manure, subsistence.agr, organic.fertilizer, native.seeds, green.manure))
```
Run the CCA and compute variance inflation factors (VIF) to check for redundancy in the predictor variables. As a rule of thumb, VIF > 10 indicates problematic levels of redundancy. Redundant predictor variables do not adversely effect model fit, but it does make it difficult to interpret the estimated effects of individual predictor variables.
```{r}
# the formula notation y ~ . means that all variables in 'fixed' should be included
my.cca <- cca(abundance.matrix2 ~ ., data=fixed)

# VIFs
vif.cca(my.cca)

```
It seems we have some problematic redundancy in our predictor variables. A principled way to choose which variable(s) to drop is to use the drop1 function, which compares the change in AIC the results from dropping each variable from the full model:
```{r}
drop1(my.cca, test="perm")
```
It looks like dropping rotation would have the least negative effect on the AIC. Let’s try dropping it and refitting the CCA:
```{r}
# drop rotation from the fixed effects
fixed <- fixed[,-2]

# rerun the cca 
my.cca <- cca(abundance.matrix2 ~ ., data=fixed)

# VIFs
vif.cca(my.cca)
```
```{r}
drop1(my.cca, test="perm")
```
Now we drop animal.traction
```{r}
# drop quelites from the fixed effects
fixed <- fixed[,-2]

# rerun the cca 
my.cca <- cca(abundance.matrix2 ~ ., data=fixed)

# VIFs
vif.cca(my.cca)
```
```{r}
drop1(my.cca, test="perm")
```
Now we drop managed.border
```{r}
# drop MixedForest from the fixed effects
fixed <- fixed[,-4]

# rerun the cca 
my.cca <- cca(abundance.matrix2 ~ ., data=fixed)

# VIFs
vif.cca(my.cca)
```
```{r}
drop1(my.cca, test="perm")
```
Now we drop organic.pesticide
```{r}
# drop MixedForest from the fixed effects
fixed <- fixed[,-9]

# rerun the cca 
my.cca <- cca(abundance.matrix2 ~ ., data=fixed)

# VIFs
vif.cca(my.cca)
```

it seems like that did the trick- all of our VIFs are now reasonable. Let’s visualize the proportion of variance explained by each CCA axis with a barplot:

```{r}
barplot(my.cca$CA$eig/my.cca$tot.chi, names.arg = 1:my.cca$CA$rank, cex.names = 0.5, ylab="Proportion of variance explained", xlab="CCA axis")
```
The first axis explains over 8% of the variance. Contrast this with the CA, for which the first axis explained more of the variation in the data. It is expected that the CCA axes will explain less variation than the CA axes because they are now constrained to be linear combinations of the predictors, whereas CA did not have any constraints.

Look at CCA plots:

```{r}
# 2 panels
#layout(matrix(1:2,1,2))

# plot sites
plot(my.cca, display=c("sites", "bp"), type="n", main="Sites", scaling="sites")
text(my.cca, display="sites", labels= practicas1$manejo,scaling="sites")
text(my.cca, display="bp", col="red")

# plot species
plot(my.cca, display=c("species", "bp"), type="n", ylab="", main="Families", scaling="species")
text(my.cca, display="species", col="black", scaling="species")
text(my.cca, display="bp", col="red")
```
As in the CA, we can see the management gradient reflected along the second axis of variation (note that the direction of the gradient is arbitrary).

Curculionidae & number of crops. Cantharidae and industrial herbicide and fertilizer.

Now extract the percent of variance explained by our predictor variables and test for a significant effect of our predictor variables on beetle community composition:

```{r}
# proportion of variance explained
my.cca$CCA$tot.chi/my.cca$tot.chi
```

```{r}
# significance test for individual predictors (type 3 test)
anova(my.cca, by="margin") 
```


```{r}
# significance test for entire model
anova(my.cca)
```

The management variables explain about 73% of varibance in beetle community composition, and the strongest predictor variables are industrial herbicide, followed by industrial pesticide, quelites, industrial fertilizer, and number of crops. The full model is statistically significant (p < 0.001).

FIN





